Demers Cartogram with Rivers

### 2023-05-11 paper/GHRVis.tex
draft: VMV 10 page
fix: grammars

\documentclass{egpubl}
\usepackage{eurovis2023}

% --- for  Annual CONFERENCE
% \ConferenceSubmission   % uncomment for Conference submission
% \ConferencePaper        % uncomment for (final) Conference Paper
% \STAR                   % uncomment for STAR contribution
% \Tutorial               % uncomment for Tutorial contribution
% \ShortPresentation      % uncomment for (final) Short Conference Presentation
% \Areas                  % uncomment for Areas contribution
% \MedicalPrize           % uncomment for Medical Prize contribution
% \Education              % uncomment for Education contribution
% \Poster                 % uncomment for Poster contribution
% \DC                     % uncomment for Doctoral Consortium
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to , special issue
\SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
%                          % EuroVis, SGP, Rendering, PG
% --- for  EG Workshop Proceedings
% \WsSubmission      % uncomment for submission to EG Workshop
% \WsPaper           % uncomment for final version of EG Workshop contribution
% \WsSubmissionJoint % for joint events, for example ICAT-EGVE
% \WsPaperJoint      % for joint events, for example ICAT-EGVE
% \Expressive        % for SBIM, CAe, NPAR
% \DigitalHeritagePaper
% \PaperL2P          % for events EG only asks for License to Publish

@@

\begin{abstract}
    Cartograms are representations of geographical and abstract data based on a value-by-area mapping technique. A relative of the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization. One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy. To address this, we propose a novel hybrid cartogram layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms. Rivers affect both the layout and appearance of cartograms. We evaluate the proposed hybrid layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset. The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram, while improving the legibility, readability, and accuracy of the cartogram.
\end{abstract}

>>>>

\documentclass{egpubl}
\usepackage{VMV2023}

% --- for  Annual CONFERENCE
% \ConferenceSubmission   % uncomment for Conference submission
% \ConferencePaper        % uncomment for (final) Conference Paper
% \STAR                   % uncomment for STAR contribution
% \Tutorial               % uncomment for Tutorial contribution
% \ShortPresentation      % uncomment for (final) Short Conference Presentation
% \Areas                  % uncomment for Areas contribution
% \MedicalPrize           % uncomment for Medical Prize contribution
% \Education              % uncomment for Education contribution
% \Poster                 % uncomment for Poster contribution
% \DC                     % uncomment for Doctoral Consortium
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to , special issue
% \SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
%                          % EuroVis, SGP, Rendering, PG
% --- for  EG Workshop Proceedings
\WsSubmission      % uncomment for submission to EG Workshop
% \WsPaper           % uncomment for final version of EG Workshop contribution
% \WsSubmissionJoint % for joint events, for example ICAT-EGVE
% \WsPaperJoint      % for joint events, for example ICAT-EGVE
% \Expressive        % for SBIM, CAe, NPAR
% \DigitalHeritagePaper
% \PaperL2P          % for events EG only asks for License to Publish

@@

\begin{abstract}
    Cartograms are representations of geographical and abstract data based on a value-by-area mapping technique. A relative of the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization. One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing legibility, readability, and geographical accuracy. To address this, we propose a novel hybrid cartogram layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms. Rivers affect both the layout and appearance of cartograms. We evaluate the proposed hybrid layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset. The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram while improving the legibility, readability, and accuracy of the cartogram.
\end{abstract}

### 2023-05-11 paper/section/data.tex
fix: grammars

Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}. The first step is to obtain both geospatial boundaries and EHR data. The second step is to pre-process the EHR data to remove empty and erroneous values. The final step is to transform the data into a format that is suitable for cartograms. Geospatial boundaries, or shapefiles, were obtained from the sources described here.

\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes over time due to NHS re-organization. The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.

\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England. These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues. While including smaller rivers is technically feasible it may not increase the legibility of the cartogram. This is an open question for future work.

@@

After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, and convert them into GeoJSON files. Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON \cite{TopoJSON} file. TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation. See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the pre-processing result. We describe the one-time pre-processing steps in more detail in \Cref{app:pre-processing}.

>>>>

Obtaining heterogeneous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}. The first step is to obtain both geospatial boundaries and EHR data. The second step is to preprocess the EHR data to remove empty and erroneous values. The final step is to transform the data into a format that is suitable for cartograms. Geospatial boundaries, or shapefiles, were obtained from the sources described here.

\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes over time due to NHS reorganization. The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG reorganizations that took place in 2021 and 2022.

\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the Thames River, the Trent River, and the Great Ouse River in England. These rivers were chosen as they are well-known rivers, pass through regions with dense populations, and provide informative geographical and topological cues. Although including smaller rivers is technically feasible, it may not increase the legibility of the cartogram. This is an open question for future work.

@@

After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust the projections and convert them into GeoJSON files. Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON file \cite{TopoJSON}. TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation. See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the preprocessing result. We describe the one-time preprocessing steps in more detail in \Cref{app:pre-processing}.

### 2023-05-11 paper/section/evaluation.tex
draft: e_t percentage
draft: review
fix: grammars

\textbf{H1.1:} The introduction of rivers can increase target location accuracy.

\textbf{H1.2:} The introduction of rivers can reduce the time needed to locate a target region in a cartogram.

We test these hypotheses using location based tasks. Given a standard choropleth on the left side of the screen and the equivalent cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right). See \Cref{fig:task}.

@@

We discuss the variables of our user study in this section.

\bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. We set this parameter to either render rivers or not.

@@

\bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram. Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants. \textit{Response time:} Another dependent variable is the time taken by the participants to complete each task.

\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms. \textit{Communicating the target CCG to the user:} We inform the participant about the target CCG that they need to find. The target CCG is shown in the form of a non-stop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.

@@

135 CCGs are rendered on the screen as a choropleth on the left. We then generated another view on the right using cartograms. The color in both visual designs is mapped to the dataset. See \Cref{fig:task} for an example.

    {
        \begin{figure}[tb!]
            \centering
            \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task.png}
            \caption{A sample location task for participants. The left shows the choropleth map, and the right shows the corresponding cartogram. Both images show the three longest rivers in England, and color is mapped to CCG population. The target CCG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram. In this figure, $ \nodeTopologicalError $ is reduced from 9 to 0. }
            \label{fig:task}
        \end{figure}
    }

@@

\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design. These tasks also serve as a quality test to see if the users are taking the study seriously. A demonstration of three sample tasks is also included in the instructional video.

\textbf{P3:} The participants are asked to complete 16 location tasks involving 4 CCGs. The response and reaction time are recorded. These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.

@@

The materials for P1-P4 are available on the Open Science Foundation's website at \URL{https://osf.io/q39w7}.

@@

In this section we analyze the results of our preliminary user study. We recruited \pCount participants to perform 16 location tasks each, in total we collected 384 responses, 24 responses that took over 60 seconds are excluded from the analysis. We hypothesize that participants taking longer than 60 seconds to complete a task were distracted from the study.

\bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants. The chart shows that the introduction of rivers has improved the accuracy for two CCGs. The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002. This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.

\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to located a CCG measured for four CCGs. The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs. The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38. However, the average time to locate a given CCG is reduced for 3 out of the 4 CCGs users in the study.

>>>>

\textbf{H1.1:} The introduction of rivers can increase the accuracy of the target location.

\textbf{H1.2:} The introduction of rivers can reduce the time required to locate a target region in a cartogram.

We test these hypotheses using location-based tasks. Given a standard choropleth on the left side of the screen and the equivalent cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right). See \Cref{fig:task}.

@@

In this section, we discuss the variables of our user study.

\bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. We set this parameter to render rivers or not.

@@

\bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram. Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants. \textit{Response time:} Another dependent variable is the time it takes the participants to complete each task.

\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms. \textit{Communicating the target CCG to the user:} We inform the participant about the target CCG they need to find. The target CCG is shown in the form of a nonstop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.

@@

135 CCGs are rendered on the screen as a choropleth on the left. We then generated another view on the right using cartograms. The color of both visual designs is mapped to the dataset. See \Cref{fig:task} for an example.

    {
        \begin{figure}[tb!]
            \centering
            \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task.png}
            \caption{A sample location task for participants. The left shows the choropleth map, and the right shows the corresponding cartogram. Both images show the three longest rivers in England, and color is mapped to CCG population. The target CCG is blinking on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram. In this figure, $ \nodeCartographicErrorMax $ = 1.875\% , and an $ \nodeTopologicalError $ of 6.67\% is eliminated. }
            \label{fig:task}
        \end{figure}
    }

@@

\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design. These tasks also serve as a quality test to see whether users are taking the study seriously. A demonstration of three sample tasks is also included in the instructional video.

\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 CCGs. Response and reaction time are recorded. These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.

@@

The materials for P1-P4 are available on the Open Science Foundation website at \URL{https://osf.io/q39w7}.

@@

In this section, we analyze the results of our preliminary user study. We recruited \pCount participants to perform 16 location tasks each, in total we collected 384 responses, 24 responses that took over 60 seconds are excluded from the analysis. We hypothesize that participants who took longer than 60 seconds to complete a task were distracted from the study.

\bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants. The chart shows that the introduction of rivers has improved the accuracy of two CCGs. The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002. This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.

\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to locate a CCG measured for four CCGs. The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs. The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38. However, the average time to locate a given CCG is reduced for 3 out of the 4 CCGs users in the study.

### 2023-05-11 paper/section/introduction.tex
fix: grammars

Among the four types of cartograms categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes. Also comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions. 
We hypothesize that introducing rivers improves the overall legibility of a cartogram. 
By \textit{legibility} we mean readability and ability to interpret the cartogram.  
We test this by asking user-study participants to perform correspondence and location tasks.
To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
We present a user study demonstrating its effectiveness.

@@

The results of the user study indicate that rivers can improve the legibility of cartograms. One of the major challenges involved is how to develop a layout algorithm that handles different shapes. In other words, the hybrid layout algorithm is novel because it handles different types of elements -- square representing regions and polylines representing rivers. Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing error.

>>>>

Among the four types of cartogram categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes. Also, comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions. 
We hypothesize that introducing rivers improves the overall legibility of a cartogram. 
By \textit{legibility} we mean readability and ability to interpret the cartogram.  
We test this by asking user-study participants to perform correspondence and location tasks.
To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
We present a user study that demonstrates its effectiveness.

@@

The results of the user study indicate that rivers can improve the legibility of cartograms. One of the major challenges involved is developing a layout algorithm that handles different shapes. In other words, the hybrid layout algorithm is novel because it handles different types of elements: square representing regions and polylines representing rivers. Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing error.

### 2023-05-11 paper/section/limitation.tex
draft: review
draft: VMV 10 page
fix: grammars

In this section we discuss some limitations of our work and future research directions. The first limitation is the color map we use to depict the data in our cartograms. We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green. However, we believe that the choice of color map can have a significant impact on the legibility of cartograms. In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target. See \Cref{fig:extreme} for an example. We plan to explore the impact of different color maps on the legibility of cartograms in future work.

{
    \begin{figure}[tb!]
        \centering
        \includegraphics[width=\columnwidth,keepaspectratio]{figure/limitations/extreme.png}
        \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
        \label{fig:extreme}
    \end{figure}
}

Another limitation is the overlap removal algorithm (FNOR) we use. We believe that developing a new algorithm with built-in constraint support can significantly reduce the time needed to generate cartograms with rivers.

>>>>

In this section, we discuss some limitations of our work and future research directions.
Because this is a new concept, it opens the door for many future research directions.
The first limitation is the color map that we use to depict the data in our cartograms. 
We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
However, we believe that the choice of color map can have a significant impact on the legibility of cartograms.
In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target.
See \Cref{fig:extreme} for an example.
We plan to explore the impact of different color maps on the legibility of cartograms in future work.

{
    \begin{figure}[tb!]
        \centering
        \includegraphics[width=0.7\columnwidth,keepaspectratio]{figure/limitations/extreme.png}
        \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
        \label{fig:extreme}
    \end{figure}
}

Another limitation is the overlap removal algorithm (FNOR) we use.
We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers. Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.

Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, coastlines, etc.
We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
We would like to explore the case of river-river intersections and testing out more geographies such as the U.S. and Europe, which have more complex rivers.
We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
Furture work also includes adding rivers to dorling-style cartograms.

### 2023-05-11 paper/section/main.tex
fix: grammars

\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries. For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel. We then load the river shapefiles and render the rivers. Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next nearest vertex. We do not need the original river resolution to incorporate them into cartograms. We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests. This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}. We further apply simplification by removing vertices that are too close to each other. The initialization procedure is a one-time process that can be saved for re-use.

\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes. We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}. Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions. An increase in $ \nodeSize $ can cause the nodes to overlap. During overlap removal, we compute node trajectories (See \Cref{alg:TranslateNode}) and translate nodes to their new position. Nodes that cross a river, denoted $ \NxR $, are translated back to their previous position. If a node oscillates across a river, we identify this as a stalemate. One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river. We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $. The gradual size increase process provides stability to the layout, as can be seen in the accompanying video.

@@

The logic for translating a node's position is detailed in \Cref{alg:TranslateNode}. When a node's position changes, we test if the node's trajectory intersects any segment of a river. See \Cref{alg:check river intersection}. A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required. Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm. We label these nodes, $ \NxR $.

@@

When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where the nodes are crossed by a translated river, denoted $ \RxN $. As a heuristic we also translate these nodes by $ \vectorAvg $. The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated in order to create space for nodes that are too crowded together. In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $. Thus, we can influence how far $ \river $ is translated in each iteration of the layout algorithm. We can use this to ensure smooth transitions between iterations.

@@

As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}. If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate the congestion. A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D). All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E). The procedure for constructing corridors is provided by \Cref{alg:derive corridor}.

@@

    {
        \begin{figure}[tb!]
            \centering
            \includegraphics[width=\columnwidth,keepaspectratio]{figure/stalemate.png}
            \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times. The node is translated back to its previous position (iteration 2). A stalemate can occur when the area is congested and the node is unable to translate to a new position without intersecting a river.}
            \label{fig:stalemate}
        \end{figure}
    }

@@

When the algorithm terminates, the node layout is considered optimal where no node has crossed, or crossed by a river (denoted by $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $). Every node remains on the same side of the river as its centroid.

@@

\Cref{fig:overview} presents an overview of the application, including the user options. The user can adjust the following parameters:

@@

\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option allows nodes to cross rivers. Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default. Both options are useful for generating different layouts and for observing the behavior of the hybrid layout algorithm.

\bobgraph{Corridor Length:} The user can define the length of a corridor, which is used to resolve stalemate situations. A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout. The default corridor length is three times the max node size.

>>>>

\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries. For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel. We then load the river shapefiles and render the rivers. Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next nearest vertex. We do not need the original river resolution to incorporate them into cartograms. We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests. This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}. We further apply simplification by removing vertices that are too close to each other. The initialization procedure is a one-time process that can be saved for reuse.

\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes. We chose FNOR over other node overlap removal algorithms because FNOR is capable of minimizing spread and node movement while maintaining a good level of global shape preservation \cite{chen2020Node}. Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions. An increase in $ \nodeSize $ can cause the nodes to overlap. During overlap removal, we compute node trajectories (See \Cref{alg:TranslateNode}) and translate nodes to their new position. Nodes that cross a river, denoted $ \NxR $, are translated back to their previous position. If a node oscillates across a river, we identify this as a stalemate. One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river. We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $. The gradual size increase process provides stability to the layout, as can be seen in the accompanying video.

@@

The logic for translating the position of a node is detailed in \Cref{alg:TranslateNode}. When a node's position changes, we test if the node's trajectory intersects any segment of a river. See \Cref{alg:check river intersection}. A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required. Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm. We label these nodes, $ \NxR $.

@@

When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where nodes are crossed by a translated river, denoted $ \RxN $. As a heuristic, we also translate these nodes by $ \vectorAvg $. The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated to create space for the nodes that are too crowded together. In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $. Thus, we can influence how far $ \river $ is translated in each iteration of the layout algorithm. We can use this to ensure smooth transitions between iterations.

@@

As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}. If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate congestion. A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D). All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E). The procedure for constructing corridors is provided by \Cref{alg:derive corridor}.

@@

    {
        \begin{figure}[tb!]
            \centering
            \includegraphics[width=\columnwidth,keepaspectratio]{figure/stalemate.png}
            \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times. The node is translated back to its previous position (iteration 2). A stalemate can occur when the area is congested and the node cannot translate to a new position without intersecting a river.}
            \label{fig:stalemate}
        \end{figure}
    }

@@

When the algorithm terminates, the node layout is considered optimal where no nodes have crossed or crossed a river (denoted $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $). Every node remains on the same side of the river as its centroid.

@@

\Cref{fig:overview} presents an overview of the application, including user options. The user can adjust the following parameters:

@@

\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option allows nodes to cross rivers. Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default. Both options are useful for generating different layouts and to observe the behavior of the hybrid layout algorithm.

\bobgraph{Corridor Length:} The user can define the length of a corridor that is used to resolve stalemate situations. A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout. The default corridor length is three times the max node size.
